---
title: "In-class Exercise 2:  Geospatial Data Wrangling with R"
date: "`r Sys.Date()`"
format: html
execute:
  message: false
  warning: false
editor: visual
---

#### Installing the R packages

First, we need to check if we have the three R packages to be used for the In-Class Excercise (sf ,tidyverse and funModelling) installed. If not, we need to install it automatically. The below code checks if the packages are already installed, else it will install the packages for us. However, before running the below code, ensure that 'packman' package has been installed in your RStudio.

```{r}
pacman::p_load(sf, tidyverse, funModeling)
```

#### Importing the Geospatial Dataset

To Import the Geospatial data, run the following code. Ensure that the directory of the files and the file name is correct.

```{r}
NGA <- st_read("data/geospatial", 
                  layer = "geoBoundaries-NGA-ADM2") %>%
  st_transform(crs = 26392)
```

#### Importing the NGA Dataset

To Import the NGAdata, run the following code. Ensure that the directory of the files and the file name is correct.

#\| code-fold: true

#\| code-summary = "Show the code"

evaluation = false

```{r}
geoNGA <- st_read("data/geospatial", 
                  layer = "nga_admbnda_adm2_osgof_20190417") %>%
  st_transform(crs = 26392)
```

Importing Aspatial data

Write the extracted data into rds format

converting aspatial data into geospatial

project transformation

https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex02/in-class_ex02
