---
title: "Take-Home Excercise 01"
author: "Mayuri Salunke"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  message: false
  warning: false
editor: visual
---

# Overview

-   Talk about water situation in Nigeria

-   Talk about water situation in Oson state

-   Talk about the Osun river and its myth

-   Objective of the assignment and elaborate

# Setup

## Packages Used

-   sf : Used for ..

-   tidyverse : Used for ..

-   tmap :

-   spatstat :

-   raster :

-   maptools :

-   rgdal :

-   

```{r}
pacman::p_load(kableExtra)
```

## Datasets Used

```{r}
# initialise a dataframe of our geospatial and aspatial dataset details
datasets <- data.frame(
  Type=c("Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         "Geospatial",
         
         "Aspatial",
         "Aspatial",
         "Aspatial",
         "Aspatial"),
  
  Name=c("Singapore National Boundary `CostalOutline`",
         "Master Plan 2014 Subzone Boundary (Web) `MP14_SUBZONE_WEB_PL`",
         "MRT & LRT Locations Aug 2021 `MRTLRTStnPtt`",
         "Bus Stop Locations Aug 2021 `BusStop`",
         "Taxi Stand Locations Aug 2021 `TaxiStop`",
         "Historic Sites",
         "Moneychanger Locations",
         "Monuments",
         "Museums",
         
         "Singapore Airbnb Listings June 2019",
         "Singapore Airbnb Listings June 2021",
         "Tourism",
         "Hotels"),
  
  Format=c(".shp", 
           ".shp", 
           ".shp", 
           ".shp", 
           ".shp", 
           ".kml",
           ".kml",
           ".kml",
           ".kml",
           
           ".csv",
           ".csv",
           ".csv",
           ".csv"),
  
  Source=c("[data.gov.sg](https://data.gov.sg/dataset/national-map-polygon)",
           "[data.gov.sg](https://data.gov.sg/dataset/master-plan-2014-subzone-boundary-web)",
           "[LTA Data Mall](https://datamall.lta.gov.sg/content/datamall/en/search_datasets.html?searchText=mrt)",
           "[LTA Data Mall](https://datamall.lta.gov.sg/content/datamall/en/search_datasets.html?searchText=bus%20stop)",
           "[LTA Data Mall](https://datamall.lta.gov.sg/content/datamall/en/search_datasets.html?searchText=taxi)",
           "[data.gov.sg](https://data.gov.sg/dataset/historic-sites)",
           "[data.gov.sg](https://data.gov.sg/dataset/locations-of-money-changer)",
           "[data.gov.sg](https://data.gov.sg/dataset/monuments)",
           "[data.gov.sg](https://data.gov.sg/dataset/museums)",
           
           "[Inside Airbnb](http://insideairbnb.com/get-the-data.html)",
           "[Inside Airbnb](http://insideairbnb.com/get-the-data.html)",
           "[OneMap API](https://www.onemap.gov.sg/docs/)",
           "[OneMap API](https://www.onemap.gov.sg/docs/)")
  )

# with reference to this guide on kableExtra:
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# kable_material is the name of the kable theme
# 'hover' for to highlight row when hovering, 'scale_down' to adjust table to fit page width
library(knitr)
library(kableExtra)
kable(datasets, caption="Datasets Used") %>%
  kable_material("hover", latex_options="scale_down")
```

# Data Wrangling : Geospatial Data

# Data Wrangling : Aspatial Data

```{r}

```

# Exploratory Spatial Data Analysis

# Specific KDE Maps

# Second-order Spatial Point Patterns Analysis

# Acknowledgments
